TacticToe: Learning the Game of Tactics

### 2018-02-05 tactictoe_jv.tex
Tweak some grammar and formatting

\begin{definition}(search tree)\\
A search tree $\mathfrak{T}$ is a sixtuple
$(\mathbb{T},\mathbb{G},\mathbb{A},T,G,A)$
that respects the following conditions:
\begin{itemize}
\item $\mathbb{T}$ is a set of tactics, $\mathbb{G}$ is a set of goals
 and $\mathbb{A}$ is a set of nodes. These sets contains only objects
 appearing in the search tree.
\item $T$ is a total function from $\mathbb{G}$ to $P(\mathbb{T})$. It takes a
goal and return the set of tactics already applied to this goal.
\item $G$ is a total function from $\mathbb{A}$ to $P(\mathbb{G})$. It takes a
node a return the set of goals in this node.

\item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to
$\mathbb{A}$. It takes a goal and a tactic and returns the node produced by the
application of the tactic to the goal, if this tactic has been applied to this
goal. Its support $Support(A)$ is by definition $\lbrace (g,t)\ |\ t \in T(g)
\rbrace$.
\item If $(g,t) \in Support (A)$, then $t(g) = G(A(g,t))$.
\end{itemize}

@@

\begin{definition}\label{def:desc}(Children, descendant, active node)\\
We now define what are the children and the descendants of a node $a$. The
children of a relative to a goal $g$ in $G(a)$
is the set of nodes produced by tactics applied to $g$:
  \[Children(a,g) = \lbrace A(g,t)\in \mathbb{A}\ |\ t \in T(g) \rbrace \]

@@

Extended children of $a$ by:
  \[ExtChildren(a) = \bigcup_{g \in G(a)} Children(a,g) \]

Descendants of $a$ by:
\begin{align*}
Descendant^{0}(a) &=_{def} \lbrace a \rbrace \\
Descendant^{n+1}(a) &=_{def} \bigcup_{a' \in Descendant^{n}(a)}
ExtChildren(a') \\
Descendant(a) &=_{def} \bigcup_{n \in \mathbb{N}} Descendant^n(a)\\
\end{align*}

@@

We now concentrate on the definition a reasonable evaluation function for the
output of tactics. Theoretically, an interesting list of goal is one
that has a short proof, and we could measure the likelihood of being
provable in a number of steps from previous proof search. However, this
approach produces a lot of negatives\todo{CK: cite millions lemma mining
or/and deep math}. So to reduce this number, we limit
ourselves to collect output of tactics tested during orthogonalization. We
declare a list of goals $l$ to be positive if it is has been produced by the
winner of an orthogonalization competition. Our predictor with radius $r$
produces a set of $r$ list of goals
closest to $l$. The
number of positive example in this set is noted $Positive(r,l)$.
During the proof search, we evaluate nodes through the list of goals they
contain using the following evaluation function:
\begin{align*}
PriorEvaluation_r (a_i) &= \frac{Positive(r,G(a_i))}{r}\\
\end{align*}

@@

\begin{definition}\label{def:value}(Value)
The value of the $i^th$ child $a_i$ of a parent $p$ is determined by:
\[Value(a_i) = CurEvaluation(a_i) + c_{exploration} * Exploration(a_i)\]

@@

\[CurEvaluation(a_i) =
  \sum_{a' \in Descendant(a_i)} \frac{PriorEvaluation(a')} {card\
  Descendant(a_i) + Failure(a_i)}\]

The number $Failure(a_i)$ is the amount of failures that occurred during node
extension from descendants of $a_i$ .

The exploration term is determined by the prior policy and the current policy.
\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{CurPolicy(a_i)}\]

\[CurPolicy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(p))}}\]
The policy is approximatively the  percentage of time a node was visited.The
square root skews this probability to favor even more exploration of nodes with
few visits.

@@

\subsection{Node extension}
Let $a$ be the result of node selection.
If $a$ is not an active part of the search then extending $a$ is useless and
the algorithm reports a failure. In our experiments, filtering nodes that not
active before node selection is
not beneficial. It may be because it breaks some symmetry in the
$Value$ function.
Otherwise, it applies the best untested tactic $t$ on the open goal $g$ of this
node according to the prediction order for $g$. If no such tactic exists or if
$t$ is not productive, a failure is returned.
Otherwise, a new node $c$ is created that contains the list of goals $l$
produced by $t$ and is added to the children of $a$.

>>>>

\begin{definition}(search tree)\\
A search tree $\mathfrak{T}$ is a sixtuple
$(\mathbb{T},\mathbb{G},\mathbb{A},T,G,A)$
that respects the following conditions:
\begin{itemize}
\item $\mathbb{T}$ is a set of tactics, $\mathbb{G}$ is a set of goals
 and $\mathbb{A}$ is a set of nodes. These sets contains only objects
 appearing in the search tree.
\item $T$ is a total function from $\mathbb{G}$ to $\mathcal{P}(\mathbb{T})$. It takes a
goal and return the set of tactics already applied to this goal.
\item $G$ is a total function from $\mathbb{A}$ to $\mathcal{P}(\mathbb{G})$. It takes a
node and returns the set of goals in this node.

\item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to
$\mathbb{A}$. It takes a goal and a tactic and returns the node produced by the
application of the tactic to the goal, if this tactic has been applied to this
goal. Its support $\mathit{Support}(A)$ is by definition $\lbrace (g,t)\ |\ t \in T(g)
\rbrace$.
\item If $(g,t) \in \mathit{Support}(A)$, then $t(g) = G(A(g,t))$.
\end{itemize}

@@

\begin{definition}\label{def:desc}(Children, descendant, active node)\\
We now define what are the children and the descendants of a node $a$. The
children of a relative to a goal $g$ in $G(a)$
is the set of nodes produced by tactics applied to $g$:
  \[\mathit{Children}(a,g) = \lbrace A(g,t)\in \mathbb{A}\ |\ t \in T(g) \rbrace \]

@@

Extended children of $a$ by:
  \[\mathit{ExtChildren}(a) = \bigcup_{g \in G(a)} \mathit{Children}(a,g) \]

Descendants of $a$ by:
\begin{align*}
\mathit{Descendant}^{0}(a) &=_{def} \lbrace a \rbrace \\
\mathit{Descendant}^{n+1}(a) &=_{def} \bigcup_{a' \in \mathit{Descendant}^{n}(a)}
\mathit{ExtChildren}(a') \\
\mathit{Descendant}(a) &=_{def} \bigcup_{n \in \mathbb{N}} \mathit{Descendant}^n(a)\\
\end{align*}

@@

We now concentrate on the definition a reasonable evaluation function for the
output of tactics. Theoretically, an interesting list of goal is one
that has a short proof, and we could measure the likelihood of being
provable in a number of steps from previous proof search. However, this
approach produces a lot of negatives\todo{CK: cite millions lemma mining
or/and deep math}. So to reduce this number, we limit
ourselves to collect output of tactics tested during orthogonalization. We
declare a list of goals $l$ to be positive if it is has been produced by the
winner of an orthogonalization competition. Our predictor with radius $r$
produces a set of $r$ list of goals
closest to $l$. The
number of positive example in this set is noted $\mathit{Positive}(r,l)$.
During the proof search, we evaluate nodes through the list of goals they
contain using the following evaluation function:
\begin{align*}
\mathit{PriorEvaluation}_r (a_i) &= \frac{\mathit{Positive}(r,G(a_i))}{r}\\
\end{align*}

@@

\begin{definition}\label{def:value}(Value)
The value of the $i^th$ child $a_i$ of a parent $p$ is determined by:
\[\mathit{Value}(a_i) = \mathit{CurEvaluation}(a_i) + c_{exploration} * \mathit{Exploration}(a_i)\]

@@

\[\mathit{CurEvaluation}(a_i) =
  \sum_{a' \in Descendant(a_i)} \frac{\mathit{PriorEvaluation}(a')} {card\
  \mathit{Descendant}(a_i) + \mathit{Failure}(a_i)}\]

The number $\mathit{Failure}(a_i)$ is the amount of failures that occurred during node
extension from descendants of $a_i$ .

The exploration term is determined by the prior policy and the current policy.
\[\mathit{Exploration}(a_i) = \frac{\mathit{PriorPolicy}(a_i)}{\mathit{CurPolicy}(a_i)}\]

\[\mathit{CurPolicy}(a_i) = \frac{1 + \mathit{Visit}(a_i))}{\sqrt{\mathit{Visit}(p))}}\]
The policy is approximatively the  percentage of time a node was visited.The
square root skews this probability to favor even more exploration of nodes with
few visits.

@@

\subsection{Node extension}
Let $a$ be the result of node selection.
If $a$ is not an active part of the search then extending $a$ is useless and
the algorithm reports a failure. In our experiments, filtering nodes that not
active before node selection is
not beneficial. It may be because it breaks some symmetry in the
$\mathit{Value}$ function.
Otherwise, it applies the best untested tactic $t$ on the open goal $g$ of this
node according to the prediction order for $g$. If no such tactic exists or if
$t$ is not productive, a failure is returned.
Otherwise, a new node $c$ is created that contains the list of goals $l$
produced by $t$ and is added to the children of $a$.
